---
title: "EBS pollock models"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    social: menu
    source_code: embed
    vertical_layout: fill
    theme: 
      version: 4
      bootswatch: minty
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(r4ss)
library(tidyverse)
library(here)
library(ggridges)

getwd()
loadup<-FALSE
#loadup<-TRUE
if (loadup){
  source(here::here("GetResults.R"))
  
  r1 <- ss_run<-SS_obj()
  ra <- SS_obj(SS_output(dir = here("ss","noramp"),verbose=FALSE),src="noramp") 
  r2 <- SS_obj(SS_output(dir = here("ss","mod")),src="mod")  
  r3 <- SS_obj(SS_output(dir = here("ss","mix")),src="mix")  
  r4 <- SS_obj(SS_output(dir = here("ss","high")),src="high")  
  #r5 <- SS_obj(SS_output(dir = here("ss","autocor")),src="autocor")  
  ss_sel <- rbind(r1$sel,r2$sel,r3$sel,r4$sel)#,r5$sel)
  #compute_matrix_summary(sel[,2:16])
  gp_run<-gp_obj()
  pm_run<-pm_obj()
  pm_run2<-pm_obj(pm2, src="Pollock_model VPA-like")
  load(here("SAM","poll23","run","model2.RData"))
  sam_run <- SAM_obj()
  am_run<-AMAK_obj(run_dir=here("amak2","runs", "base"))
  am_run2<-AMAK_obj(run_dir=here("amak2","runs", "3par"),src="AMAK-3par-logistic")
  am_run3<-AMAK_obj(run_dir=here("amak2","runs", "cpue"),src="AMAK w CPUE", nind=6)
  all_sel <- rbind(sam_run$sel,pm_run$sel,am_run$sel,ss_run$sel,gp_run$sel)
  all_ts <- rbind( sam_run$ts,pm_run$ts,am_run$ts,ss_run$ts,gp_run$ts)
  save.image(file=here::here("compares.Rdata"))
} else {
  load(here::here("compares.Rdata"))
}

```

Platforms
=======================================================================

### Run descriptions

Each platform was intended to include as much of the configuration
and baseline data from the pollock model as possible.

  - **SS**: Stock synthesis 3

  - **GOA_model**: Uses the model as specified for the Gulf of Alaska--fewer ages included, time-varying parametric selectivity

  - **SAM**: Anders Nielsen's state-space implementation

  - **AMAK**: Updated assessment model for Alaska

  - **Pollock_model**: Base customized model original developed from 1997 (evolved over the years)
  

### Selectivity

```{r}
p<-Plot_Sel() + ggthemes::theme_few(base_size=9)
p
```


### Selectivity by age

```{r}
p <- Plot_Sel_age()
ggplotly(p)
```

### SSB

```{r}

p<- Plot_SSB()
ggplotly(p)

```

### Stock recruitment relationship

```{r}

p<- Plot_SRR()
ggplotly(p)

```
AMAK runs
=======================================================================

### Run description

Runs with different selectivity assumptions where:     

  - **base**: selectivity at age allowed to vary (sigma penalty=0.7)     
  
  - **cpue**: As base but with the early CPUE data included
  
  - **dbl_logistic**: selectivity at age with TV selectivity parameters (3-parameter logistic)

```{r like_table, echo=FALSE}
# Convert list of lists to tibble
df<-  cbind(am_run$lst[grep("Like_Comp",names(am_run$lst))[2]] ,
       as_tibble(am_run$lst[grep("Like_Comp",names(am_run$lst))[1]]),
       as_tibble(am_run2$lst[grep("Like_Comp",names(am_run2$lst))[1]]),
       as_tibble(am_run3$lst[grep("Like_Comp",names(am_run3$lst))[1]])
       )
names(df) <- c("NLL Component", "base","dbl_logistic","cpue")
df <- df |> rowwise() |> mutate(across(2:4, ~ . - min(c_across(2:4))))

gt::gt(df) |> gt::fmt_number( columns = 2:4, decimals = 1)
```

### Selectivity

```{r echo=FALSE}
# am_run<-AMAK_obj()
  #source(here::here("GetResults.R"))
#am_run2<-AMAK_obj(run_dir=here("amak2","runs", "3par"))
 # am_run<-AMAK_obj(run_dir=here("amak2","runs", "base"))
 # am_run2<-AMAK_obj(run_dir=here("amak2","runs", "3par"))
  #am_run3<-AMAK_obj(run_dir=here("amak2","runs", "cpue"),src="AMAK w CPUE", nind=6)
   am_sel <- rbind(pm_run$sel, pm_run2$sel, am_run$sel,am_run2$sel, am_run3$sel)
   am_ts <- rbind(pm_run$ts, pm_run2$ts, am_run$ts,am_run2$ts,am_run3$ts)
   am_fit<- rbind(am_run$fit, am_run2$fit, am_run3$fit)
p1 <- Plot_Sel(am_sel)
p1 + ggthemes::theme_few(base_size=9)
```

### Selectivity at age
```{r echo=FALSE}
p1  <- Plot_Sel_age(am_sel)
ggplotly(p1)
```

### SSB and recruitment

```{r echo=FALSE}
p1 <- Plot_SSB(am_ts)
ggplotly(p1)
```

### AMAK fit to indices

#### All 
```{r echo=FALSE}
p1 <- Plot_index(df=am_fit, idx=c(1,2,5)) + ggthemes::theme_few(base_size=9)
ggplotly(p1)
```

#### With CPUE 
```{r idxe2, echo=FALSE}
#glimpse(am_run3$fit)
Plot_index(df=am_run3$fit, idx=c(1,2,5,6)) + ggthemes::theme_few(base_size=9)
```

### Stock recruitment relationship

```{r}
p<- Plot_SRR(df=am_ts)
ggplotly(p)

```

SS runs
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------
### Run description

Runs with different selectivity assumptions where:     
  - **base**: selectivity at age allowed to vary (sigma penalty=0.7)     
  - **high**: selectivity at age constrained (sigma penalty=0.05)     
  - **mod**: selectivity at age moderately constrained (sigma penalty=0.4)     
  - **mix**: selectivity at age moderately constrained for middle ages, high for older ages, loose for younger ages      

### Selectivity

```{r echo=FALSE}
ss_sel <- rbind(r1$sel,r2$sel,r3$sel,r4$sel)#,r5$sel)
p1 <- Plot_Sel(ss_sel)
p1
```


### Selectivity at age
```{r echo=FALSE}
p1  <- Plot_Sel_age(ss_sel)
ggplotly(p1)
```

### SSB and recruitment

```{r echo=FALSE}
ss_ts  <- rbind(r1$ts,r2$ts,r3$ts,r4$ts)#,r5$sel)
p1 <- Plot_SSB(ss_ts)
ggplotly(p1)
```


With and w/o ramp
=======================================================================


### SSB and recruitment

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Get results w/ and w/o  ramp
df <- rbind(r1$ts,ra$ts)
  p1  <- df |> filter(Year<2024,Year>1953) |> ggplot(aes( x=Year, y= value,color=source)) + 
    geom_line(stat='identity') + geom_point(stat='identity') +
    ggthemes::theme_few() + ylab("SSB") + xlab("Year") +
    ylim(0,NA) + facet_grid(type~.,scales="free_y")  
  ggplotly(p1)
  
```

