---
output: html_document
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 90
---

```{r startup, echo=TRUE, warnings=FALSE, message=FALSE, eval=TRUE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(ebswp)
library(tidyverse)
.OVERLAY=TRUE
.THEME=ggthemes::theme_few()
options(warn=-1)
```

# September evaluations

## Drop the early CPUE data

Given the SSC's request, we re-ran the selected 2023 model where these data are ignored.
The results are shown in the table below.

```{r dropCPUE}
#| echo: TRUE
#| warnings: FALSE
#| messages: FALSE
#| label: fig-dropCPUE
#| fig.cap: 'SRR curves as estimated in the 2023 assessment (top) and conditioned on an alternative Fmsy assumptions (middle and bottom).'
library(patchwork)
modlst <- list()
 modlst[[1]] <- read_rep(here::here("2024", "runs", "lastyr", "pm.rep"))
 modlst[[2]] <- read_rep(here::here("2024", "runs", "dropCPUE_sept", "pm.rep"))
 modlst[[3]] <- read_rep(here::here("2024", "runs", "srr_int_sept", "pm.rep"))
 modlst[[4]] <- read_rep(here::here("2024", "runs", "SST", "pm.rep"))
 modlst[[5]] <- read_rep(here::here("2024", "runs", "SSTshrt", "pm.rep"))
 
p1<- plot_srr(modlst[c(1,2)], alpha = .2, xlim = c(0, 6400), ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021)) + ggtitle("SRR as estimated in the 2023 assessment")
n

```
# Including temperature effects in the stock-recruitment relationship


## Tier 1 and ecosystem function evaluation

The SSC requested an evaluation of the ecosystem function as part of the SRR consideration
and Tier 1 control rules within the FMP. Their comment was:

*"The SSC would prefer not to
make a risk table adjustment based on the difference from Tier 1 to Tier 3 again during
the 2024 assessment cycle. The SSC requests that the next stock assessment bring back a
new approach that may include development of a constant buffer based on factors extrinsic
to the stock assessment (ecosystem function), or a better representation of the
uncertainty in the Tier 1 and control rule calculations such that a reduction from maximum
ABC is not needed every year."*

Our approach to this was to evaluate the ecosystem function by comparing the current model
with a model where the MSY value is set to 1.75 and 1.3 Mt. The results are shown in the
figure below.

### Inverting the MSY value to show the stock-recruitment relationship (SRR)

The ecosystem function is evaluated by comparing the current model with a model where the
MSY value is set to 1.75 and 1.3 Mt (@fig-srrplot_cond). When overplotted, the fit
comparisons indicate somewhat more degradations for conditioning the curve such that MSY
equals 1.75 and 1.3 Mt (@fig-srrplot_cond2).

As part of the SRR conditioning exercise, where the MSY value is a function of the
selectivity estimate, we also evaluated the family of SRR curves given the past 20 years
of selectivity estimates (the base result used the mean selectivity over the recent 2
years; specifically, the mean selectivity for 2021 and 2022 for the 2023 terminal year
assessment).

```{r srrplot_cond}
#| echo: TRUE
#| warnings: FALSE
#| messages: FALSE
#| label: fig-srrplot_cond
#| fig.width: 7
#| fig.height: 8
#| fig.cap: 'SRR curves as estimated in the 2023 assessment (top) and conditioned on an alternative Fmsy assumptions (middle and bottom).'
library(patchwork)
 srlst <- list()
 srlst[[1]] <- read_rep(here::here("2024", "runs", "lastyr", "pm.rep"))
 srlst[[2]] <- read_rep(here::here("2024", "runs", "condSRR_sept", "MSY175.rep"))
 srlst[[3]] <- read_rep(here::here("2024", "runs", "condSRR_sept", "MSY130.rep"))
names(srlst) <- c("Last year", "MSY=1.75 Mt", "MSY=1.3 Mt")
p1<- plot_srr(srlst[c(1)], alpha = .2, xlim = c(0, 6400), ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021)) + ggtitle("SRR as estimated in the 2023 assessment")
p2<- plot_srr(srlst[c(2)], alpha = .2, xlim = c(0, 6400), ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021)) + ggtitle("SRR condition to have MSY=1.75 Mt")
p3<- plot_srr(srlst[c(3)], alpha = .2, xlim = c(0, 6400), ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021)) + ggtitle("SRR condition to have MSY=1.3 Mt")
p1/p2/p3 + plot_layout(axis_titles = "collect")

```

```{r srrplot_cond2}
#| echo: TRUE
#| warnings: FALSE
#| messages: FALSE
#| label: fig-srrplot_cond2
#| fig.cap: 'SRR curves as estimated in the 2023 assessment (top) and conditioned on an alternative Fmsy assumptions (middle and bottom).'
### How sensitive is the curve to the selectivity assumption

  p1 <- plot_srr(srlst[c(1:3)], alpha = .2, xlim = c(0, 6400), ylim = c(0, 85000), 
              sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021)) + ggtitle("SRR as estimated in the 2023 assessment")
p1
#plotly::ggplotly(p1)
```

For this set of experiments, we evaluated the SRR curve given the past 20 years of
selectivity estimates. From this we get a set of SRR curves that are conditioned on the
selectivity estimate AND the MSY value. The script to run this set involved reading in the
control file and modifying the option for which years to include for the selectivity
estimate (can be specific years or a range of years from which to use).

Initial results showed that conditioning the curve to make the MSY value 1.3 Mt was
insensitive to the selectivity estimate (@fig-selsrrplot). Presumably this was due to the
scale parameter (i.e., R_0) being adjusted to get the MSY value to 1.3 Mt.

We evaluated the ABC value for each of the 20 years of selectivity estimates both for the
case where the SRR productivity was assumed to be 1.3 Mt and for the case where this
constraint was removed. This showed a high range of variability for ABC as a function of
the selectivity assumption (@fig-sel_ABC_plot).

```{r selsrrplot}
#| echo: TRUE
#| warnings: FALSE
#| message: FALSE
#| eval: TRUE
#| label: fig-selsrrplot
#| fig.cap: 'SRR curves as estimated in the 2023 assessment (top) and conditioned on an alternative Fmsy assumptions (middle and bottom).'

modlst <- list()
for (i in 3:20) {
  modlst[[i-2]] <- read_rep(here::here("2024", "runs", "condSRR_sept", paste0("sel_", i, ".rep")))
}
names(modlst) <- paste0("sel_", (3):20)
#names(modlst)
#.OVERLAY=TRUE
plot_srr(modlst[c(1,4,12,18)], alpha = .2, xlim = c(0, 6400), ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021))

```


```{r sel_ABC_plot, echo=TRUE, warnings=FALSE, message=FALSE, eval=TRUE}
#| echo: TRUE
#| warnings: FALSE
#| message: FALSE
#| eval: TRUE
#| label: fig-sel_ABC_plot
#| fig.cap: 'SRR curves as estimated in the 2023 assessment (top) and conditioned on an alternative Fmsy assumptions (middle and bottom).'
dftmp <- data.frame()
modlst <- list()
for (i in 3:20) {
  modlst[[i-2]] <- read_rep(here::here("2024", "runs", "selsens_sept", paste0("sel_", i, ".rep")))
}
names(modlst) <- paste0("sel_", (3):20)
#names(modlst)
#.OVERLAY=TRUE
#plot_srr(modlst[c(1,18)], alpha = .2, xlim = c(0, 6400), ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021))

dftmp <- data.frame()
for (i in 1:18) {
  dftmp <- rbind(dftmp, data.frame(
    sel_yr = 2023 - i-2,
    ABC = modlst[[i]]$T1[1, 2],
    NLL = modlst[[i]]$tot_like,
    NLL_data = modlst[[i]]$dat_like,
    NLL_rec = sum(modlst[[i]]$rec_like),
    NLL_pri = sum(modlst[[i]]$Priors)
  ))
}
M <- modlst[[1]]
# glimpse(dftmp)
(dftmp) |> ggplot(aes(x = sel_yr, y = ABC, color = sel_yr)) +
  geom_point() +
  ggthemes::theme_few() + ylab("ABC, unconstrained ") + xlab("Selectivity year") +
  ggtitle("SRR conditioning on selectivity data") +
  theme(legend.position = "none")

```

### Simulation testing an alternative management procedure

EBS pollock "maintain ecosystem function" catch-advice rule as requested by the SSC. One
idea would be to evaluate the role of pollock as a key part of the forage base (say of 1-3
yr old pollock). A management goal with an explicit consideration of ecosystem function
would be to avoid low levels. For example, if the forage base appeared to be close to say
the lower 20th percentile from historical estimates, then a management procedure might
include an adjustment that would occur then to avoid any further declines
(@fig-quant_prey).

Practical aspects of such a management procedure might be prohibitive since in a
projection scenario, information on those age groups would be limited. Consequently, we
evaluated the prey base compared to spawning biomass (@fig-ssb_prey). This figure shows
that the relationship is poorly determined but that the forage component does tend to
decline at lower spawning biomass levels. This suggests that by monitoring spawning
biomass, there is potential linkages to downstream impacts of the main forage ages.

The fisheries management plan (FMP) for the BSAI is subject to the categorization of the
stock assessment to obtain the maximum permissible ABC and OFL. For pollock, this falls on
the determination of the appropriateness of the $F_{MSY}$ and the probability distribution
of that value (uncertainty). This relates directly to a number of key factors, including
the selectivity, the SRR, and future weight-at-age ($F_{MSY}$ applies to numbers of fish,
but ABC is in biomass). within the FMP, an ABC can be set below the maximum permissible
value. Without making an explicit ammendment to the FMP, we propose an alternative
semi-empirical approach. The first principle of this approach is to maintain the ecosystem
function by contrasting from what's been observed over the last several decades. Going
forwardd, the catch advice could be adjusted based on the spawning biomass relative to the
historical mean. This would be a simple rule that would adjust the catch advice based on
the spawning biomass relative to the historical mean. I.e., if the catch in the current
year is say 1.2 million t, and the SSB next year is 30% above the mean, then with a
regulator to dampen change, next year’s recommendation would be
$1.2 \times \sqrt{1.30} = 1.368$ million t. Similarly, if the SSB next year was only 75%
of the mean value, the recommendation would be $1.2 \times \sqrt{0.75} = 1.039$ million t.
If the SSB stayed at 75% of the mean, then the following year would be
$1.039 \times \sqrt{0.75} = 08998$ million t.

Some rationale for this strategy might include

-   The detrimental effects of fishing on the ecosystem appear to be mainly
    sustainable...are there assignable issues from fishing that affected the ecosystem
    function that we’ve measured with any degree of confidence?

-   If there hadn’t been any pollock fishing ever, would we be able to note that the
    ecosystem would be in a better state?

-   Tuning the adjustments based on historical catches and ecosystem outcomes would
    generally be deemed "acceptable"

For the latter part of the above, we could use the historical data on diet composition in
a mass-balance ecosystem model to evaluate the effects of the catch advice on the
ecosystem. This was pursued in the next section.

```{r quant_prey}
#| label: fig-quant_prey
#| fig.cap: "Historical age-1 to age-3 pollock abundance as estimated from the assessment model."
#| warnings: FALSE
#| message: FALSE
#| 
M <- read_rep(here::here("2024","runs","test","pm.rep"))
df <- data.frame(Year = 1964:2023, forage = rowSums(M$N), SSB = M$SSB[, 2]) |> mutate(rel_forage = forage / mean(forage), rel_SSB = SSB / mean(SSB))
qfor <- quantile(df$forage, c(0.05, .2))
df |> ggplot(aes(x = Year, y = forage)) +
  geom_line() +
  labs(title = "Total pollock forage (age 1-3)", x = "Year", y = "Abundance (thousands)") +
  ggthemes::theme_few() +
  xlim(c(1980, 2024)) +
  geom_hline(yintercept = qfor[1], color = "red") +
  geom_hline(yintercept = qfor[2], color = "blue", type = 2)
```

```{r ssb_prey}
#| warnings: FALSE
#| message: FALSE
#| label: fig-ssb_prey
#| fig.cap: "Historical spawning biomass and 'prey' abundance for pollock as estimated from the assessment model."
df |>
  pivot_longer(cols = 4:5, names_to = "var", values_to = "value") |> filter(var=="rel_forage") |> 
  ggplot(aes(x = log(SSB), y = log(forage) )) + geom_point() +
  geom_smooth() +
  labs(title = "EBS Pollock ", x = "SSB", y = "Forage (ages 1 to 3 abundance)") +
  ggthemes::theme_few() #+
  #xlim(c(1980, 2024))

```

```{r ssb_hist}
#| warnings: FALSE
#| message: FALSE
#| label: fig-ssb_hist
#| fig.cap: "Historical spawning biomass relative to the mean for pollock as estimated 
#| from the assessment model. Red horizontal line is the mean value."
df |> ggplot(aes(x = Year, y = SSB)) +
  geom_line() +
  labs(title = "Spawning biomass of pollock ", x = "Year", y = "Biomass (t) ") +
  ggthemes::theme_few() +
  xlim(c(1980, 2024)) +
  geom_hline(yintercept = 2257, color = "red")

```

```{r mp_hist}
#| warnings: FALSE
#| message: FALSE
#| label: fig-mp_hist
#| fig.cap: "Different levels of proposed ABCs given spawning biomass running means (ssb_mn) along
#| with projected spawning biomass (ssb_next) and actual historical catches (Current_catch). 
#| The panels represent alternative values of responsiveness ($/lambda$)."
#| 

df1 <- NULL
for (resp in c(0.05,0.5,0.95)){
  for (i in 2000:2022) {
    # get values from the time series
    lastidx <- i - 1977 + 14
    cat_mn <- mean(M$obs_catch[14:lastidx])
    cat_this <- M$obs_catch[lastidx]
    ssb_mn <- mean(M$SSB[14:lastidx, 2])
    ssb_this <- (M$SSB[lastidx, 2])
    ssb_next <- (M$SSB[lastidx + 1, 2])
    # cat_next   <- (ssb_next/ssb_this)^0.5 * cat_this
    cat_next <- (ssb_next / ssb_mn)^resp * cat_this
    df1 <- data.frame(Year = i, Mean_catch = cat_mn, Current_catch = cat_this, ABC = cat_next, ssb_mn = ssb_mn, ssb_this = ssb_this, ssb_next = ssb_next, resp=resp) |> rbind(df1)
  }
}

df1 |>
  pivot_longer(cols = c(3:5, 7), names_to = "var", values_to = "value") |>
  mutate(Year = ifelse(var == "ABC", Year + 1, Year)) |>
  ggplot(aes(x = Year, y = value, shape = var, color = var)) +
  geom_point() +
  geom_line() +
  labs(title = "EBS Pollock ", x = "Year", y = "relative to mean") +
  theme_bw() + facet_grid(resp~.) +
  xlim(c(2000, 2024))
```

Such a management procedure would need to be tested using a full feedback loop system.
Consequently, we updated the simulation feature of the assessment model to behave as an
operating model. This required some simplifications on how the bottom-trawl survey
covariance structure was applied, in addition to a couple of other issues (e.g., the
random-effects derived estimates of weight-at-age).

### Tests of future ecosystem states using EWE (Rpath) scenarios

To develop a more empirically based management procedure for pollock we recognize that
this species is a keystone component for the Bering Sea. We thus posed the question on,
given available understanding of ecosystem linkages, what might things look like if the
fishery intensified or if it was removed altogether and compare that with "status quo"
fishing.

The EBS pollock model has contained code to serve as an operating model for simulating
datasets and be used as a self testing platform. Several modifications that have
occurred since this was first implemented, including the addition of a new biomass index, 
required some code updates. 

The model was run for 100 years under three scenarios: status




## Profiling and sensitivity of the variability specified about the SRR

The SSC requested that the full time series be included for the SRR conditioning.

Additionally, they requested that we ensure that the bias correction is applied.

We confirm that in past assessments, for the period which the SRR curve was applied, the
bias correction term was included. Specifically,

$$ \hat{R}_t = f(B_{t-1}) e^{\epsilon_t-0.5 \sigma_R^2} $$

where $\epsilon_t \sim N(0, \sigma_R^2)$. Here the production function (which generates
age-1 recruitment) is a function of spawning biomass in the previous year ($f(B_{t-1})$).
Since this function "generates" age-1 recruitment from a lognormal distribution, the basis
for this must be scaled accordingly. Therefore the assessment model numbers-at-age one
($\dot{R_t}=N_{1,t}$), must account for the bias based on the SRR. This leads to the
recruitment component of the negative log-likelihood as

$$ -ln(L_{rec}) = \sum_{t=1}^{T} \left[\frac{\left(\chi_t + \frac{\sigma^2_R}{2}\right)^2}{2\sigma^2_R} + ln(\sigma_R) \right] $$

where $\chi_t = \log(\dot{R}_t) - \log(\hat{R}_t)$. Note that the bias correction term
falls within the likelihood because the bias is a function of the model estimates.

```{r sigmaRprofile2, eval=FALSE, echo=TRUE}
#| echo: TRUE
#| warnings: FALSE
#| message: FALSE
#| eval: FALSE

ctl.orig <- read_dat(here::here("2024", "runs", "sigr", "control.dat"))
ctl <- ctl.orig
ctl$phase_sigr
ctl$sigrprior
setwd(here::here("2024", "runs", "sigr"))
for (i in seq(.53, .83, 0.02)) {
  ctl$sigrprior <- i
  write_dat(output_file = "control.dat", ctl)
  system("./pm -iprint 300 ")
  file.copy(from = "pm.rep", to = paste0("sigr_", i, ".rep"))
}

```

```{r sigmaRplot1}
#| echo: TRUE
#| warnings: FALSE
#| message: FALSE
#| eval: TRUE
#| label: fig-sigmRplot1
#| fig.cap: 'SRR curves as estimated in the 2023 assessment (top) and conditioned on an alternative Fmsy assumptions (middle and bottom).'

#---Now read in results and plot-------------------------
#modlst[[i]] <- read_rep(paste0("sigr_",i,".rep"))
ii=seq(.53,1.05,0.02) 
modlst <- list()
for (i in  1:27 ) {
  modlst[[i]] <- read_rep(here::here("2024","runs","sigr",paste0("sigr_",ii[i],".rep")))  
}
names(modlst) <- paste0("sigr_",ii)
df_sig <- NULL
for (i in  1:27 ) {
  df_sig <- rbind(df_sig, data.frame( 
    sigmaR=modlst[[i]]$sigr,
    NLL_rec=sum(modlst[[i]]$rec_like),
    NLL_pri=sum(modlst[[i]]$Priors),
    NLL=modlst[[i]]$tot_like,
    ABC=modlst[[i]]$T1[1,2],
    NLL_data = modlst[[i]]$dat_like
  )
  )
}
# glimpse(df_sig)
df_sig |> mutate(NLL = NLL - min(NLL), priors=NLL_pri-min(NLL_pri), 
                 NLL_rec = NLL_rec - min(NLL_rec), NLL_data = NLL_data - min(NLL_data)) |> 
  pivot_longer(cols = c("NLL","NLL_rec","priors", "NLL_data"), names_to = "type", values_to = "NLL") |>
  ggplot(aes(x=sigmaR,y=NLL, color=type)) + geom_line(linewidth=1.4) + ggthemes::theme_few() +
  coord_cartesian( ylim = c(0, 4))
```

```{r sigmaR_abc}
#| echo: TRUE
#| eval: TRUE
#| warnings: FALSE
#| message: FALSE
#| label: fig-sigmR_abc
#| fig.cap: 'Impact of $\sigma_R$ on the ABC value from the 2023 assessment.'

df_sig |> 
  ggplot(aes(x=sigmaR,y=ABC, color="red")) + geom_line(linewidth=1.4) + ggthemes::theme_few()

#plot_sigr(modlst, alpha = .2, xlim = c(0,4200), ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021))  

#setwd(here::here())
plot_srr(modlst[c(1,11,21)], alpha = .2, xlim = c(0,4200),  ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021))
```

```{r sigrplot2}
#| echo: TRUE
#| eval: TRUE
#| warnings: FALSE
#| message: FALSE
#| label: fig-sigrplot2
#| fig.cap: "SRR curves as estimated in the 2023 assessment for different fixed values of sigmaR." 

plot_srr(modlst[c(1,6,11,16,21,26)], alpha = .2, xlim = c(0,4200), 
         ylim = c(0, 85000), sizeout = 2, sizein = 4, yrsin = c(1977, 1979:2021))
```
